Learning Network Size while Training with ShrinkNets

### 2018-02-09 paper/ICML/approach.tex
edits

In this section we describe the ShrinkNets approach, which allows us to learn
the network size as part of the training process. The key idea is the addition
of \emph{switch layers} added after each layer in the network;  these layers
allow us to disable certain neurons during training.  Once we've introducd the
basic method, we explain how to adapt the training procedure to support this new
layer.

@@

At a high-level, the ShrinkNets approach consists of two interconnected stages.
The first stage identifies neurons that do not contribute to increasing the
prediction accuracy of the network and deacivates them. The second stage is
concerned with actually removing the neurons from the network and shrinking its
size, therefore leading to faster inference times. We give an overview of both
stages next:

\noindent\textbf{Deactivating Neurons On-The-Fly: }During the first stage,
ShrinkNets applies an on/off switch to every neuron of an initially oversized
network. We model the on/off switches by multiplying each input (or output) of
each layer by a parameter $\bm{\theta}$, with values $0$ or $1$. A value of $0$
will deactivate the neuron, while $1$ will let the signal go through. These
switches are part of a new layer, called \textbf{switch layer}, which can
be applied to fully connected as well as convolutional layers.

We want to minimize the number of on switches to reduce the model size as much
as we can, while preserving prediction accuracy. This can be achieved by jointly
minimizing the objective of the network and a factor of the L0 norm of the
vector containing all the on/off switches. Because finding an optimal binary
assignement is an NP-Hard problem, we allow $\bm{\theta}$ to be a real number
instead of a 0/1 value, thus constraining the L1 norm as opposed to the L0 norm.

\noindent\textbf{Neuron Removal: } During this stage, the neurons that are
deactivated by the switch layers are actually removed from the network,
effectively shrinking the network size. This is what leads to faster inference
times, as we demonstrate in our evaluation. We choose to remove neurons at
training time because we have observed that allows the active neurons to adapt
to the new network architecture. Existing techniques focus on neuron removal
after training, and require an extra fine-tuning process to compensate for the
removal.

In the remainder of this section we describe in detail the switch layer as well
as how to adapt the training process for ShrinkNets.

@@

Let $L$ be a layer in a neural network that takes an input tensor $\bm{x}$ and
produces an output tensor $y$ of shape $\left(c \times d_1 \times \dots \times
d_n\right)$ where $c$ is the number of neurons in that layer.  For instance, for
fully connected layers, $n$=0 and the output is single dimensional of size $c$
(ignoring batch size for now) while for a 2-D convolutional layer, $n$=2 and $c$
is the number of output channels or feature maps.

We want to tune the size of $L$ by applying a switch layer, $S$. The switch
layer is parametrized by a vector $\theta \in \mathbb{R}^c$ such that the result
of applying $S$ to $L(\bm{x})$ is a tensor also of size $\left(c \times d_1
\times \dots \times d_n\right)$ such that:

%Suppose we wish to tune the size of $L$ by applying a switch layer.
%A switch layer $S$ applied to the output of $L$ can be parametrized by a 
%vector

\begin{equation} 
S_{\theta}(L(\bm{x}))_{i,...} = \bm{\theta}_iL(\bm{x})_{i, ...}
\end{equation}

Effectively, once passed through the switch layer, each output channel $i$
produced by $L$ is scaled by the corresponding $\theta_i$. Note that when
$\theta_i = 0$, the $i^{\text{th}}$ channel is multiplied by zero and will not
contribute to any computations after the switch layer. If this happens, we say
the switch layer has {\it deactivated} the neuron of layer $L$ corresponding to
channel $i$.

@@

%Formally, 
For training, we need to account for the effect of the swith layers in the loss
function. The effect of switch layers can be expressed in terms of a sparsity constraint
that pushes values in the $\theta$ vector to 0. Then, given a neural network
parameterized by weights $\bm{W}$ and switch layer parameters $\bm{\theta}$, we
optimize the following ShrinkNet loss:

\begin{equation}
  L_{SN}(\bm{x},\bm{y};\bm{W}, \bm{\theta}) = L(\bm{x}, \bm{y}; \bm{W}) +
  \lambda\norm{\theta}_1 + \lambda_2\norm{\bm{W}}_p
\end{equation}

i.e., the expression augments the regular training loss with a regularization
term for the switch parameters and another on the network weights.

>>>>

In this section we describe the ShrinkNets approach, which allows us to learn
the network size as part of the training process. The key idea is the addition
of \emph{switch layers} added after each layer in the network;  these layers
allow us to disable certain neurons during training.  We start by introducing the
basic method, and then explain how to adapt the training procedure to support this new
layer.

@@

At a high-level, out approach consists of two interconnected stages.
The first stage identifies neurons that do not improve the
prediction accuracy of the network and deactivates them. The second stage then
removes neurons from the network, shrinking its
size and yielding faster inference times. Broadly, these stages work as follows:

\noindent\textbf{Deactivating Neurons On-The-Fly: }During the first stage,
ShrinkNets applies an on/off switch to every neuron of an initially over-sized
network. We model the on/off switches by multiplying each input (or output) of
each layer by a parameter $\bm{\theta}\in\{0,1\}$. A value of $0$
will deactivate the neuron, while $1$ will let the signal go through. These
switches are part of a new layer, called the \textbf{switch layer}; this layer applies 
to fully connected as well as convolutional layers.

Our objective is to minimize the number of ``on'' switches to reduce the model size as much
as possible while preserving prediction accuracy. This can be achieved by jointly
minimizing the objective of the network and a factor of the L0 norm of the
vector containing all the switches. Because finding an optimal binary
assignment is an NP-Hard problem, we relax the problem by allowing
 $\bm{\theta}$ to be a real number
instead of a binary value and constrain it using the L1 instead of L0 norm.

\noindent\textbf{Neuron Removal: } During this stage, the neurons that are
deactivated by the switch layers are actually removed from the network,
effectively shrinking the network size. This step improves inference
times, as we demonstrate in our evaluation. We choose to remove neurons at
training time because we have observed that this allows the remaining active neurons to adapt
to the new network architecture. Existing techniques focus on neuron removal
after training, and require an extra fine-tuning process to compensate for the
removal.\srm{cite something?}

In the remainder of this section we describe in detail the switch layer as well
as and the training process for ShrinkNets, and then describe the removal process in Section~\ref{neuron_killing}.

@@

Let $L$ be a layer in a neural network that takes an input tensor $\bm{x}$ and
produces an output tensor $y$ of shape $\left(c \times d_1 \times \dots \times
d_n\right)$ where $c$ is the number of neurons in that layer.  For instance, for
fully connected layers, $n$=0 and the output is single dimensional vector of size $c$
(ignoring batch size for now) while for a 2-D convolutional layer, $n$=2 and $c$
is the number of output channels or feature maps.

We want to tune the size of $L$ by applying a switch layer, $S$, containing $c$
switches.  The switch
layer is parametrized by a vector $\theta \in \mathbb{R}^c$ such that the result
of applying $S$ to $L(\bm{x})$ is a also a tensor size $\left(c \times d_1
\times \dots \times d_n\right)$ such that: 
%Suppose we wish to tune the size of $L$ by applying a switch layer.
%A switch layer $S$ applied to the output of $L$ can be parametrized by a 
%vector 
\begin{equation} 
S_{\theta}(L(\bm{x}))_{i,...} = \bm{\theta}_iL(\bm{x})_{i, ...} \forall i \in [1\ldots c]
\end{equation}
Once passed through the switch layer, each output channel $i$
produced by $L$ is scaled by the corresponding $\theta_i$. Note that when
$\theta_i = 0$, the $i^{\text{th}}$ channel is multiplied by zero and will not
contribute to any computation after the switch layer. If this happens, we say
the switch layer has {\it deactivated} the neuron corresponding to channel $i$ of layer $L$.

@@

%Formally, 
For training, we need to account for the effect of the switch layers on the loss
function. The effect of switch layers can be expressed in terms of a sparsity constraint
that pushes values in the $\theta$ vector to 0. In this way, given a neural network
parameterized by weights $\bm{W}$ and switch layer parameters $\bm{\theta}$, we
optimize ShrinkNet loss as follows 
\begin{equation}
\small
  L_{SN}(\bm{x},\bm{y};\bm{W}, \bm{\theta}) = L(\bm{x}, \bm{y}; \bm{W}) +
  \lambda\norm{\theta}_1 + \lambda_2\norm{\bm{W}}_p
\end{equation}
This expression augments the regular training loss with a regularization
term for the switch parameters and another on the network weights.
